---
output:
  word_document: default
  html_document: default
---

### Question 1

#### Part a
##### Decrement table
Need a vector of ages x and the two vectors of independent rates, q_x^d and q_x^w.
```{r}
x<-c()
x[55:58]<-c(55:58)

qx<-c()
qx[55:57]<-c(0.001046,0.001199,0.001375)
```

Choose the radix as $l_0=10000$, use the formulae $l_{x+1}=l_x*(1-q_x)$.
```{r}
lx<-c()
lx[55]<-10000

for (j in 55:58){
  lx[j+1]<-lx[j]*(1-qx[j])
}
```
Print all the decrement detail.
```{r}
haha<-cbind(x[55:58],lx[55:58],qx[55:58])
colnames(haha)<-c("x","lx","qx")
haha
```

we can calculate the policy values using tabulated annuity and assurance functions as below.$$_tV=Sa_{x+t\,\overline{n-t}|}\,@4\,\%$$ The effective interest is 4% because the benefits are increasing geometrically.$$0.04=\frac{1.06}{1.0192308}-1$$

We can use recursive formulae or use the relation between annuity and assurance.

$${a}_{x+t\,\overline{n-t}|}=vp_{x}(1+{a}_{x+t+1\,\overline{n-t-1}|})$$
$${a}_{x+t\,\overline{n-t}|}=\frac{v-{A}_{x+t\,\overline{n-t}|}}{d}$$

```{r}
px<-c()
px[55:57]<-1-qx[55:57]

a.PV<-c()
v<-1.0192308/1.06
a.PV[57]<-1*v*px[57]
a.PV[58]<-0

for (j in 57:56){
  a.PV[j-1]<-v*px[j-1]*(1+a.PV[j])
}
cbind(x[55:57],px[55:57],a.PV[55:57])
```

#### In-force cash flows
For this three year temporary annuity policy, we need the vectors of
1.	Premia
2.	Expenses
3.	Benefits
4.	Policy values

```{r}
t<-c(1:3)
prems.BoY<-c(40250,0,0)

initial.expenses.BoY<-c()
initial.expenses.BoY[1]<-100+0.015*prems.BoY[1]
initial.expenses.BoY[2:3]<-c(0,0)


pay.expenses.BoY<-c()
expenses.inflation<-1.05^(0:2)
pay.expenses.BoY[1:3]<-10*expenses.inflation

benefits.EoY<-15000*(1.0192308)^c(0:2)


policy.EoY<-c()
policy.EoY[55:57]<-benefits.EoY[1:3]*a.PV[56:58]


wawa<-cbind(t,prems.BoY[1:3],initial.expenses.BoY[1:3],pay.expenses.BoY[1:3],benefits.EoY[1:3],policy.EoY[55:57])
colnames(wawa)<-c("year","prems.BoY","initial.expenses.BoY","pay.expenses.BoY",
                 "benefits.EoY","policy.EoY")
wawa
```

#### Expected amounts per sold policy
Here we need vectors of
1.	expected premium
2.	expected expenses
3.  Benefits
4.	expected transfer 
5.	expected interest earned
6.	expected policy value 

```{r}
Eprem.BoY <- c()
for (j in 1:3){
  Eprem.BoY[j]<-prems.BoY[j]*lx[j+54]/lx[55]
}
Eprem.BoY

```

```{r}
Eexp1.BoY <- c()
for (j in 1:3){
  Eexp1.BoY[j]<-initial.expenses.BoY[j]*lx[j+54]/lx[55]
}
Eexp1.BoY
```

```{r}
Eexp2.BoY <- c()
for (j in 1:3){
  Eexp2.BoY[j]<-pay.expenses.BoY[j]*lx[j+55]/lx[55]
}
Eexp2.BoY
```


```{r}
Ebft.EoY <- c()
for (j in 55:57){
  Ebft.EoY[j]<-benefits.EoY[j-54]*lx[j+1]/lx[55]
}
Ebft.EoY 
```

```{r}
Epolval.EoY <- c()
for (j in 55:57){
  Epolval.EoY[j]<-policy.EoY[j]*lx[j+1]/lx[55]
}
Epolval.EoY 
```

```{r}
i<-0.08

Einterest <- c()
Einterest[1] <- (Eprem.BoY[1]-Eexp1.BoY[1])*i
Einterest[2:3] <- (Eprem.BoY[2:3]-Eexp1.BoY[2:3]+Epolval.EoY[55:56])*i
Einterest


EPolValChg <- c()
EPolValChg[55] <- Epolval.EoY[55]
EPolValChg[56:57] <- Epolval.EoY[56:57]-Epolval.EoY[55:56]
EPolValChg
```

The profit signature shows as below
```{r}
ETransfer <-Eprem.BoY-Eexp1.BoY-Eexp2.BoY-Ebft.EoY+Einterest-EPolValChg
ETransfer
```


the expected net present value of profits is 201.6858 and the profit margin is 0.005010827.
```{r}
EPVPremsER <- sum(Eprem.BoY/(1.12)^(0:2))
EPVPremsER

EPVTransfersER <- sum(ETransfer[55:57]/(1.12)^(1:3))
EPVTransfersER 

```

```{r}
ProfMargER <- EPVTransfersER/EPVPremsER
ProfMargER
```

#### part b
the yield on transfers is 0.5028818.
```{r}
q1<-function(i){sum(ETransfer[55:57]/(1+i)^(1:3))}
uniroot(q1,c(0.4,0.6))
```


