---
output:
  word_document: default
  html_document: default
---

### Q2
#### Independent rates
```{r}
x<-c()
x[30:40]<-c(30:40)

qx.d.ind<-c()
qx.d.ind[30:39]<-c(0.00043,0.00042,0.00041,0.0004,0.00041,
                   0.00042,0.00043,0.00045,0.00049,0.00053)

qx.w.ind<-c()
qx.w.ind[30:39]<-c(0.250,0.200,0.175,rep(0.150,6),1)

```

#### Dependent rates
We can now calculate the multiple decrement table. first for deaths.
```{r}
qx.d.dep<-c()
qx.d.dep[30:39]<-qx.d.ind[30:39]*(1-0.5*qx.w.ind[30:39])

```

And then for withdrawls.
```{r}
qx.w.dep <- c()
qx.w.dep[30:39] <- qx.w.ind[30:39]*(1-0.5*qx.d.ind[30:39])
```

#### Double decrement table

```{r}
al.x <- c()
al.x[30] <- 100000
```

initialise $(ad)_x^d$ and $(ad)_x^w$


```{r}
ad.x.d <- c()
ad.x.w <- c()

for (j in 30:39) {
  ad.x.d[j] <- al.x[j]*qx.d.dep[j]
  ad.x.w[j] <- al.x[j]*qx.w.dep[j]
  al.x[j+1] <- al.x[j]-ad.x.d[j]-ad.x.w[j]
}

```

print the results
```{r}
lala<-cbind(x[30:40],qx.d.ind[30:40],qx.w.ind[30:40],
            qx.d.dep[30:40],qx.w.dep[30:40],al.x[30:40],ad.x.d[30:40],ad.x.w[30:40])

colnames(lala)<-c("x","qx.d.ind","qx.w.ind","qx.d.dep","qx.w.dep","al.x","ad.d.x","ad.w.x")
lala

```

### Unit Fund

#### calculating the possible amounts for an in-force policy
```{r}
at<-c(0.45,rep(0.01,9))
at
b<-c(rep(0.005,10))
b
p<-5000

inv.prems.BoY<-c()
inv.prems.BoY<-p*(1-at)*(1-b)
inv.prems.BoY

m<-0.003
VoU.EoY<-c()
VoU.EoY[1]<-inv.prems.BoY[1]*1.1*(1-m)
for (j in 1:9){
  VoU.EoY[j+1]<-(inv.prems.BoY[j+1]+VoU.EoY[j])*1.1*(1-m)
}
VoU.EoY
```

#### Expected amounts per initial policy
```{r}
Einv.prem.BoY<-c()
for (j in 1:10){
  Einv.prem.BoY[j]<-inv.prems.BoY[j]*al.x[30+j-1]/al.x[30]
}

EVoU.EoY<-c()
for (j in 1:10){
  EVoU.EoY[j]<-VoU.EoY[j]*al.x[30+j]/al.x[30]
}

Eint.UL.EoY<-c()
Eint.UL.EoY[1]<-Einv.prem.BoY[1]*0.1
for (j in 2:10){
  Eint.UL.EoY[j]<-(Einv.prem.BoY[j]+EVoU.EoY[j-1])*0.1
}

transfer.UL<-c()
transfer.UL[1]<-Einv.prem.BoY[1]+Eint.UL.EoY[1]-EVoU.EoY[1]
for (j in 2:10){
  transfer.UL[j]<-Einv.prem.BoY[j]+Eint.UL.EoY[j]+EVoU.EoY[j-1]-EVoU.EoY[j]
}
cbind(Einv.prem.BoY,Eint.UL.EoY,EVoU.EoY,transfer.UL)

```

#### Non-unit fund
FOR AN IN-FORCE POLICY
```{r}
premfee.BoY<-at*p

margins.BoY<-b*(p-premfee.BoY)

expenses.NUL.EoY<-c()
expenses.NUL.EoY[1]<-0.45*p+58
expenses.NUL.EoY[2:10]<-58*1.2^c(1:9)

Bft.NUL.EoY<-VoU.EoY

cbind(premfee.BoY,margins.BoY,expenses.NUL.EoY,Bft.NUL.EoY)
```

EXPECTED AMOUNTS PER INITIAL POLICY
BOY premium based fees, buy-sell margins & expenses
```{r}
Epremfee.BoY<-c()
Emargins.BoY<-c()
Eexpenses.NUL.EoY<-c()

for (j in 1:10){
  Epremfee.BoY[j]<-premfee.BoY[j]*al.x[29+j]/al.x[30]
  Emargins.BoY[j]<-margins.BoY[j]*al.x[29+j]/al.x[30]
  Eexpenses.NUL.EoY[j]<-expenses.NUL.EoY[j]*al.x[29+j]/al.x[30]
}

Epremfee.BoY
Emargins.BoY
Eexpenses.NUL.EoY

```

```{r}
Edeath.Bft.EoY<-c()
Esurr.Bft.EoY<-c() 
for (j in 1:10){
  Edeath.Bft.EoY[j]<-Bft.NUL.EoY[j]*ad.x.d[29+j]/al.x[30]
  Esurr.Bft.EoY[j]<-Bft.NUL.EoY[j]*ad.x.w[29+j]/al.x[30]
}
Edeath.Bft.EoY  
Esurr.Bft.EoY  
```

```{r}
Eint.NUL.EoY<-(Epremfee.BoY+Emargins.BoY-Eexpenses.NUL.EoY)*0.08
Eint.NUL.EoY

transfer.NUL<-Epremfee.BoY+Emargins.BoY-Eexpenses.NUL.EoY+Eint.NUL.EoY+
              transfer.UL-Edeath.Bft.EoY-Esurr.Bft.EoY 
transfer.NUL



```
profit margin
```{r}

Etransfer.NUL<-sum(transfer.NUL*1.125^-c(1:10))
Etransfer.NUL

Etransfer.NUL/p

```

#### part b

```{r}
q2<-function(s){
  x<-c()
  x[30:40]<-c(30:40)
  
  qx.d.ind<-c()
  qx.d.ind[30:39]<-c(0.00043,0.00042,0.00041,0.0004,0.00041,
                     0.00042,0.00043,0.00045,0.00049,0.00053)
  
  qx.w.ind<-c()
  qx.w.ind[30:39]<-c(0.250,0.200,0.175,rep(0.150,6),1)
  
  qx.d.dep<-c()
  qx.d.dep[30:39]<-qx.d.ind[30:39]*(1-0.5*qx.w.ind[30:39])
  
  qx.w.dep <- c()
  qx.w.dep[30:39] <- qx.w.ind[30:39]*(1-0.5*qx.d.ind[30:39])
  
  
  al.x <- c()
  al.x[30] <- 100000
  
  ad.x.d <- c()
  ad.x.w <- c()
  
  for (j in 30:39) {
    ad.x.d[j] <- al.x[j]*qx.d.dep[j]
    ad.x.w[j] <- al.x[j]*qx.w.dep[j]
    al.x[j+1] <- al.x[j]-ad.x.d[j]-ad.x.w[j]
  }
  
  at<-c(0.45,rep(0.01,9))
  
  b<-c(rep(0.005,10))
  
  p<-5000
  
  inv.prems.BoY<-c()
  inv.prems.BoY<-p*(1-at)*(1-b)
  
  
  m<-0.003
  VoU.EoY<-c()
  VoU.EoY[1]<-inv.prems.BoY[1]*1.1*(1-m)
  for (j in 1:9){
    VoU.EoY[j+1]<-(inv.prems.BoY[j+1]+VoU.EoY[j])*1.1*(1-m)
  }
  VoU.EoY[9]<-s
  
  Einv.prem.BoY<-c()
  for (j in 1:10){
    Einv.prem.BoY[j]<-inv.prems.BoY[j]*al.x[30+j-1]/al.x[30]
  }
  
  
  EVoU.EoY<-c()
  for (j in 1:10){
    EVoU.EoY[j]<-VoU.EoY[j]*al.x[30+j]/al.x[30]
  }
  
  Eint.UL.EoY<-c()
  Eint.UL.EoY[1]<-Einv.prem.BoY[1]*0.1
  for (j in 2:10){
    Eint.UL.EoY[j]<-(Einv.prem.BoY[j]+EVoU.EoY[j-1])*0.1
  }
  
  transfer.UL<-c()
  transfer.UL[1]<-Einv.prem.BoY[1]+Eint.UL.EoY[1]-EVoU.EoY[1]
  for (j in 2:10){
    transfer.UL[j]<-Einv.prem.BoY[j]+Eint.UL.EoY[j]+EVoU.EoY[j-1]-EVoU.EoY[j]
  }
  
  premfee.BoY<-at*p
  
  margins.BoY<-b*(p-premfee.BoY)
  
  expenses.NUL.EoY<-c()
  expenses.NUL.EoY[1]<-0.45*p+58
  expenses.NUL.EoY[2:10]<-58*1.2^c(1:9)
  
  Bft.NUL.EoY<-VoU.EoY
  
  Epremfee.BoY<-c()
  Emargins.BoY<-c()
  Eexpenses.NUL.EoY<-c()
  
  for (j in 1:10){
    Epremfee.BoY[j]<-premfee.BoY[j]*al.x[29+j]/al.x[30]
    Emargins.BoY[j]<-margins.BoY[j]*al.x[29+j]/al.x[30]
    Eexpenses.NUL.EoY[j]<-expenses.NUL.EoY[j]*al.x[29+j]/al.x[30]
  }
  
  Edeath.Bft.EoY<-c()
  Esurr.Bft.EoY<-c() 
  for (j in 1:10){
    Edeath.Bft.EoY[j]<-Bft.NUL.EoY[j]*ad.x.d[29+j]/al.x[30]
    Esurr.Bft.EoY[j]<-Bft.NUL.EoY[j]*ad.x.w[29+j]/al.x[30]
  }
  
  Eint.NUL.EoY<-(Epremfee.BoY+Emargins.BoY-Eexpenses.NUL.EoY)*0.08
  
  transfer.NUL<-Epremfee.BoY+Emargins.BoY-Eexpenses.NUL.EoY+Eint.NUL.EoY+
    transfer.UL-Edeath.Bft.EoY-Esurr.Bft.EoY 
  transfer.NUL[10]
  
}

uniroot(q2,c(60000,70000))
```


#### part c
Decreased. Money has been held back in the Unit Fund at time 9 (rather than being released to shareholders) to give a zero transfer at time 10. Since these additional funds earn only 10% p.a. interest in the Unit Fund, whereas the discount rate is 12.50%, overall the business is now (slightly) less profitable.
